Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-13 resex.tex
Revised results on 7th-turn pulsing. Intermediate step in revision of figures for this section. Defined "nth-turn pulsing" commands. Moved figures from appendix to text.

@@

\subsection{Pulsing every $7^{\mathrm{th}}$ turn}
\label{sec:simex7}

\begin{figure*}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png}
	\end{minipage}	
	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (top left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (top right), only in H (bottom left) and only in V (bottom right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295). The excitation is 96~nrad in both planes. The $7Q_x$ and $7Q_y$ resonance are both excited and in addition to the 14th order $7Q_x+7Q_y$.}
\end{figure*}

\begin{table}
	\caption{\label{tab:7thexp}%
		Overview of the effect of $7^{\mathrm{th}}$~turn pulsing during the two resonant excitation experiments in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}. The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V for horizontal and vertical at the same time.
	}
	\begin{ruledtabular}
		\begin{tabular}{lcc}
			Parameter & Experiment 2017 & Experiment 2016  \\
			\colrule
			optics & 2017 injection optics & 2016 injection optics \\
			tune $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\\hline
			excitation plane  & H,V and H+V&  H\\
			losses & similar loss rates for  & large losses for \\
			& pulsing in H,V &  pulsing in H \\
			&  and H+V & \\
			hor. emit. growth & no & yes, very small \\
			vert. emit. growth & yes, for pulsing in & no\\
			&  V and H+V & \\
		\end{tabular}
	\end{ruledtabular}
\end{table}

\begin{figure*}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png}
	\end{minipage}
	\caption{\label{fig:7thsim} Relative intensity (top left), relative vertical emittance (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$. The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing with two different excitation amplitudes plus a random dipole noise component in H+V of 6~nrad.}
\end{figure*}

\begin{figure*}
		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
		\centering
		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
	\end{minipage}	
		2017 experiment, $7^\mathrm{th}$ H+V\\
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
	\end{minipage}	
	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude. For all bunches the damper was not active.}
\end{figure*}

@@

@@

The experiments were divided into different periods: the first period
without excitation, the second period with a resonant excitation with
a maximum amplitude of
$A_{\mathrm{max}} = 5 \, \Delta A = \q{6}{nrad}$, and a third period
in which the excitation amplitude was further increased to
$A_{\mathrm{max}} = \q{12}{nrad}$. In~2016, experiments included a
fourth period at $A_{\mathrm{max}} = \q{24}{nrad}$.

The machine tunes were changed in standard operations from (64.28,
59.31) in~2016 to (62.27, 60.295) in~2017. This change was accompanied
by a small change in optics, considered negligible for these
measurements. A direct comparison of the two experiments is therefore
not possible. However, these differences provide a rough estimate of
how the effects of resonant excitations are affected by lattice
changes.

An insight on how the change in tune entails a change in driving
resonances is provided by frequency-map analysis.
Fig.~\ref{fig:patternfma} (bottom left and bottom right plots), which
was described previously, shows a strong increase in diffusion for
$7^{\mathrm{th}}$~turn pulsing in the horizontal plane around the
$7Q_x$ resonances, and only small changes for pulsing in the vertical
plane. In the case of the 2017 optics and tune, both the $7Q_x$ and
the $7Q_y$ resonances cross the tune footprint
(Fig.~\ref{fig:7th2017fma}, top left). Both of these resonances are
excited by $7^{\mathrm{th}}$~turn pulsing, as one can see from the
increase in tune jitter around the corresponding resonance lines for
each excitation plane separately, and thus also for their combined
action (Fig.~\ref{fig:7th2017fma}, bottom left, bottom right, and top
right plots). Although much weaker, an increase in tune diffusion is
also observed around the $7Q_x + 7Q_y$ line when pulsing in the
vertical plane or in both planes (diagonal yellow-green line in
Fig.~\ref{fig:7th2017fma}, top right and bottom right plots). This
suggests an excitation of this $14^{\mathrm{th}}$-order resonance. The
reason why this phenomenon is much less pronounced for horizontal
excitations is not understood.

The simulation results from the distribution tracking shown in Fig.~\ref{fig:7thsim} in general predict in terms of emittance for the two experiments:
\begin{description}
	\item[2016 experiment] Excitation amplitude dependent emittance growth in the horizontal but not in the vertical plane.
	\item[2017 experiment] Excitation amplitude dependent emittance growth in the vertical plane but not in the horizontal plane for an excitation in V and H+V. For an excitation only in H no emittance growth is observed.
\end{description}
Similar changes of the emittance as predicted in simulations were also
experimentally observed as depicted in Fig.~\ref{fig:7thexp}, where
emittance growth was observed only in the horizontal plane for pulsing
in H in 2016, and only in the vertical plane for pulsing in V and H+V
in 2017 with no emittance growth at all for pulsing in H. Just as for
the $10^{\mathrm{th}}$~turn pulsing, the emittance growth can be
divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast
adjustment of the beam distribution and the reach of a new equilibrium
state indicated in blue (fast adjustment) and black (equilibrium
state) in Fig.~\ref{fig:7thexp}. Also in this case, the distribution
changes are directly visible on the beam profiles taken with the BSRT
(see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexpprof}).

The losses measured during the experiment however differ from the
simulations in terms of for which case the largest rate is observed as
well as quantitatively for the specific cases. For example smaller
losses are expected for the 2016 experiments compared to the 2017
experiments from simulations, while experimentally higher loss rates
were measured in 2016 (see Fig.~\ref{fig:7thexploss}). For the 2017
experiment, the highest loss rates are expected for pulsing in H+V,
than in H and then with a considerably reduction in V. During the
experiment comparable loss rates are however measured for all three
planes (see Appendix~\ref{app:sec:7},
Fig.~\ref{fig:7thexploss2017}). Last but not least the loss rates as
defined in Eqn.~\ref{eqn:lossrate} for the 2016 (red and yellow) and
2017 (black and gree) experiments are compared in
Fig.~\ref{fig:7thexploss}. In both cases and as also observed for the
$10^{\mathrm{th}}$~turn pulsing, the loss rate follows a quadratic
scaling with the excitation amplitude. Loss rates were in general
higher during the 2016 experiments. The loss rates from simulations
are not shown in Fig.~\ref{fig:7thexploss} as losses were only
observed for much higher excitation amplitudes than tried in the
experiments.

@@

\clearpage

@@

\clearpage

\subsection{Effect of the transverse damper\label{sec:damp}}
\begin{figure*}[b]
	\centering
	damper not active\\
	\begin{minipage}[t]{0.32\linewidth}
		\centering
	random excitation V, 2017\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_no_damper_no_text.png}
	\end{minipage}	
	damper active\\
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		random excitation V, 2017 exp.\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_with_damper_no_text.png}
	\end{minipage}	
	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for a random excitation only in V (left) and for $7^{\mathrm{th}}$ (center) and $10^{\mathrm{th}}$~turn pulsing only in V (right).}
\end{figure*}
For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not damping any resonant excitation. On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation. As an example, the vertical emittance growth for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in V is compared with a random excitation in V in Fig.~\ref{fig:damp}. The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the $7^{\mathrm{th}}$~turn pulsing in H+V with a random excitation in H+V. For the $7^{\mathrm{th}}$~turn pulsing the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible. Until present, it is unknown why the transverse damper (ADT) appears to be incapable of damping any resonant excitation as it fulfills all requirements to detect as well as damp the caused oscillation:
\begin{itemize}
	\item The closed orbit distortion caused by the resonant excitation during the experiments is large enough to be detectable with the ADT pickups.
	\item The ADT compares the position of each bunch with its position in the previous turn in order to detect an orbit distortion to be damped. Therefore an oscillation due to any $k^{\mathrm{th}}$~turn pulsing with $k>1$ would be noticed and damped.
	\item The ADT is capable of damping the oscillation of each bunch individually. The ADT should therefore be capable of applying the correct corrective kick for each group of bunches seeing the same excitation amplitude.
\end{itemize}
Further studies are therefore needed in order to obtain a deeper understanding of the interaction of the transverse damper with a resonant excitation.
\clearpage

@@

\newpage

@@

\newpage

\section{Additional plots for $7^{\mathrm{th}}$~turn pulsing during 2017 experiments}
\label{app:sec:7}
\begin{figure}[h]
	\begin{minipage}[t]{1.0\linewidth}
	\centering
	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
	\end{minipage}
	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment. The excitation amplitude error is 15\% indicated by horizontal error bars. The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
\end{figure}
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		reference bunch,\\ no excitation
		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
	\end{minipage}
	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments. The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V. For both bunches the transverse damper is not active. The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
\end{figure}

\newpage

>>>>

\newcommand{\seventhtp}{7th-turn pulsing}
\newcommand{\eighthtp}{8th-turn pulsing}
\newcommand{\tenthtp}{10th-turn pulsing}

@@

\subsection{Pulsing every 7th turn}
\label{sec:simex7}

\begin{figure*}
  \begin{tabular}{cc}
    no excitation & 7th, H+V \\
    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png} &
    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png} \\
    7th, H & 7th, V \\
    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
  \end{tabular}
  \caption{FMA in betatron tune space based on the 2017 injection
    optics with no machine errors and tunes (62.27, 60.295): without
    excitation (top left) and for \seventhtp\ in H+V (top right), H
    only (bottom left) and V only (bottom right). The excitation
    amplitude is 96~nrad. The $7 Q_x$, $7 Q_y$, and 14th-order
    $7 Q_x + 7 Q_y$ resonances are excited.}
  \label{fig:7th2017fma} 
\end{figure*}

\begin{figure*}
  \begin{tabular}{cc}
    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png} &
    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png} \\
    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png} &
    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png} \\
  \end{tabular}
  \caption{Calculated bunch intensities and emittances from
    distribution tracking based on the 2016 injection optics with
    standard errors and $(Q_x, Q_y) = (64.28, 59.31)$ and 2017
    injection optics with standard errors and
    $(Q_x, Q_y) = (62.27, 60.295)$: relative intensity (top left);
    horizontal emittance (bottom left); vertical emittance (bottom
    right); the relative vertical emittance is also shown (top
    right). The solid black line includes only a random dipole noise
    component in H+V of 6~nrad. The dotted and dashed lines correspond
    to \seventhtp\ with two different excitation amplitudes (24~nrad
    and 48~nrad), plus a random dipole noise component in H+V of
    6~nrad.}
  \label{fig:7thsim}
\end{figure*}

\begin{figure*}
  \begin{tabular}{ccc}
    \multicolumn{3}{c}{2016 experiment, 7th, H}	\\
    \includegraphics[width=0.32\textwidth]{2016_bunch_intensity_h7th_no_damper_avg.png} &
    \includegraphics[width=0.32\textwidth]{2016_emith_avg_rel_h7th_no_damper.png} &
    \includegraphics[width=0.32\textwidth]{2016_emitv_avg_rel_h7th_no_damper.png} \\
    \multicolumn{3}{c}{2017 experiment, 7th, H+V} \\
    \includegraphics[width=0.32\textwidth]{2017_bunch_intensity_hv7th_no_damper_avg.png} &
    \includegraphics[width=0.32\textwidth]{2017_emith_avg_rel_hv7th_no_damper.png} &
    \includegraphics[width=0.32\textwidth]{2017_emitv_avg_rel_hv7th_no_damper.png} \\
  \end{tabular}
  \caption{Measured losses and emittances during the 2016 and 2017
    experiments: relative losses (left), relative horizontal
    emittances (middle), and relative vertical emittances
    (right). Measurements are averaged over the bunches experiencing
    the same excitation amplitude. The transverse damping system was
    not active in this set of measurements.}
  \label{fig:7thexp}
\end{figure*}

% \begin{figure*}
% 		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
% 		\centering
% 		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
% 	\end{minipage}	
% 	\begin{minipage}[t]{0.32\linewidth}
% 		\centering
% 		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
% 	\end{minipage}	
% 	\begin{minipage}[t]{0.32\linewidth}
% 		\centering
% 		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
% 	\end{minipage}	
% 		2017 experiment, $7^\mathrm{th}$ H+V\\
% 	\begin{minipage}[t]{0.32\linewidth}
% 		\centering
% 		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
% 	\end{minipage}	
% 	\begin{minipage}[t]{0.32\linewidth}
% 		\centering
% 		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
% 	\end{minipage}	
% 	\begin{minipage}[t]{0.32\linewidth}
% 		\centering
% 		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
% 	\end{minipage}	
% 	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude. For all bunches the damper was not active.}
% \end{figure*}

@@

\begin{figure}
	\begin{minipage}[t]{1.0\linewidth}
	\centering
	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
	\end{minipage}
	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment. The excitation amplitude error is 15\% indicated by horizontal error bars. The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
      \end{figure}

\begin{figure}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		reference bunch,\\ no excitation
		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
	\end{minipage}
	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments. The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V. For both bunches the transverse damper is not active. The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
\end{figure}

@@

The experiments were divided into different periods: the first period
without excitation; the second period with a resonant excitation with
a maximum amplitude of
$A_{\mathrm{max}} = 5 \, \Delta A = \q{6}{nrad}$; and a third period
in which the excitation amplitude was further increased to
$A_{\mathrm{max}} = \q{12}{nrad}$. In~2016, experiments included a
fourth period at $A_{\mathrm{max}} = \q{24}{nrad}$.

The machine tunes were changed in standard operations from (64.28,
59.31) in~2016 to (62.27, 60.295) in~2017. This change was accompanied
by a small change in optics, considered negligible for these
measurements. A direct comparison of the two experiments is therefore
not possible. However, these differences provide an estimate of how
the effects of resonant excitations are affected by lattice changes.

An insight on how the change in tune entails a change in driving
resonances is provided by frequency-map analysis.
Figure~\ref{fig:patternfma} (bottom left and bottom right plots),
which was described previously, shows a strong increase in diffusion
for $7^{\mathrm{th}}$~turn pulsing in the horizontal plane around the
$7Q_x$ resonances, and only small changes for pulsing in the vertical
plane. In the case of the 2017 optics and tune, both the $7Q_x$ and
the $7Q_y$ resonances cross the tune footprint
(Fig.~\ref{fig:7th2017fma}, top left). Both of these resonances are
excited by $7^{\mathrm{th}}$~turn pulsing, as one can see from the
increase in tune jitter around the corresponding resonance lines for
each excitation plane separately, and thus also for their combined
action (Fig.~\ref{fig:7th2017fma}, bottom left, bottom right, and top
right plots). Although much weaker, an increase in tune diffusion is
also observed around the $7Q_x + 7Q_y$ line when pulsing in the
vertical plane or in both planes (diagonal yellow-green line in
Fig.~\ref{fig:7th2017fma}, top right and bottom right plots). This
suggests an excitation of the $14^{\mathrm{th}}$-order resonance. The
reason why this phenomenon is much less pronounced for horizontal
excitations is not understood.

The simulated losses and emittances for both 2016 and 2017 conditions
are shown in Fig.~\ref{fig:7thsim}. The experimental results are
summarized in Figures~\ref{fig:7thexp}, \ref{fig:7thexploss},
\ref{fig:7thexploss2017}, and~\ref{fig:7thexpprof}.

Calculated losses for~2016 and~2017 are similar at 24~nrad and
increase more rapidly with amplitude for 2017 conditions
(Fig.~\ref{fig:7thsim}, top left). Horizontal pulsing is a few times
stronger than vertical pulsing, and the combined H+V action appears
slightly more effective than the sum of H and V separately.

Experimental losses are plotted as a function of time in
Fig.~\ref{fig:7thexp} and as a function of excitation amplitude in
Figs.~\ref{fig:7thexploss}
and~\ref{fig:7thexploss2017}. Figure~\ref{fig:7thexploss} shows that
measured losses were actually larger in~2016 than they were
in~2017. They were also a few times larger than those predicted by
simulations. Another discrepancy was observed in the excitation plane:
horizontal, vertical, and combined pulsing had similar effects
(Fig.~\ref{fig:7thexploss2017}). The same considerations on systematic
effects discussed in Section~\ref{sec:simex10} apply in this case.

This pulsing pattern generated stronger losses than expected. In this
case, we conclude that predictions are difficult, due to their
sensitivity to lattice configuration, noise sources, and beam
distributions.

The simulated emittances from distribution tracking are shown in
Fig.~\ref{fig:7thsim} (bottom left and right). The main calculated
results are the following, for the two experimental conditions:
%
\begin{description}
\item[2016 conditions] In this case, excitations were only in the
  horizontal plane. There is a fast increase of the horizontal
  emittance, dependent on the excitation amplitude, after which a
  constant emittance growth is observed. Vertical emittance growth is
  almost negligible.
\item[2017 conditions] Horizontal emittances are mostly affected by
  the amplitude of H+V excitations and, to a much smaller extent, by
  H~excitations; no effect is observed from V excitations. Vertical
  emittances are mostly affected by the amplitude of H+V and V
  excitations; the effect of H excitations is much smaller, if any,
  with no clear dependence on amplitude.
\end{description}

A set of measured emittances is plotted in Fig.~\ref{fig:7thexp}
(middle and right plots). In~2016, with only horizontal excitations,
amplitude-dependent horizontal emittance growth was observed, with no
effect on the vertical emittance. In~2017, V and H+V excitations
generated vertical emittance growth, while horizontal excitations had
no effect. These observations are consistent with the predicted
behavior.

During these experiments, emittances changed in two phases, as they
did during the $10^{\mathrm{th}}$~turn pulsing studies
(Section~\ref{sec:simex10}) and in simulations: a fast adjustment
phase (shaded in blue in Fig.~\ref{fig:7thexp}) followed by a new
equilibrium (in gray). Changes in the transverse beam distribution
could be observed directly through changes in the synchrotron-light
(BSRT) profiles (Fig.~\ref{fig:7thexpprof}).

@@

@@

\subsection{Effect of the transverse damper\label{sec:damp}}
\begin{figure*}[b]
	\centering
	damper not active\\
	\begin{minipage}[t]{0.32\linewidth}
		\centering
	random excitation V, 2017\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_no_damper_no_text.png}
	\end{minipage}	
	damper active\\
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		random excitation V, 2017 exp.\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
	\end{minipage}	
	\begin{minipage}[t]{0.32\linewidth}
		\centering
		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_with_damper_no_text.png}
	\end{minipage}	
	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for a random excitation only in V (left) and for $7^{\mathrm{th}}$ (center) and $10^{\mathrm{th}}$~turn pulsing only in V (right).}
\end{figure*}
For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not damping any resonant excitation. On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation. As an example, the vertical emittance growth for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in V is compared with a random excitation in V in Fig.~\ref{fig:damp}. The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the $7^{\mathrm{th}}$~turn pulsing in H+V with a random excitation in H+V. For the $7^{\mathrm{th}}$~turn pulsing the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible. Until present, it is unknown why the transverse damper (ADT) appears to be incapable of damping any resonant excitation as it fulfills all requirements to detect as well as damp the caused oscillation:
\begin{itemize}
	\item The closed orbit distortion caused by the resonant excitation during the experiments is large enough to be detectable with the ADT pickups.
	\item The ADT compares the position of each bunch with its position in the previous turn in order to detect an orbit distortion to be damped. Therefore an oscillation due to any $k^{\mathrm{th}}$~turn pulsing with $k>1$ would be noticed and damped.
	\item The ADT is capable of damping the oscillation of each bunch individually. The ADT should therefore be capable of applying the correct corrective kick for each group of bunches seeing the same excitation amplitude.
\end{itemize}
Further studies are therefore needed in order to obtain a deeper understanding of the interaction of the transverse damper with a resonant excitation.

@@

@@

% \section{Additional plots for $7^{\mathrm{th}}$~turn pulsing during 2017 experiments}
% \label{app:sec:7}
% \begin{figure}[h]
% 	\begin{minipage}[t]{1.0\linewidth}
% 	\centering
% 	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
% 	\end{minipage}
% 	\begin{minipage}[t]{1.0\linewidth}
% 		\centering
% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
% 	\end{minipage}
% 	\begin{minipage}[t]{1.0\linewidth}
% 		\centering
% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
% 	\end{minipage}
% 	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment. The excitation amplitude error is 15\% indicated by horizontal error bars. The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
% \end{figure}
% \begin{figure}[h]
% 	\begin{minipage}[t]{0.49\linewidth}
% 		\centering
% 		reference bunch,\\ no excitation
% 		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
% 	\end{minipage}
% 	\begin{minipage}[t]{0.49\linewidth}
% 		\centering
% 		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
% 		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
% 	\end{minipage}
% 	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments. The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V. For both bunches the transverse damper is not active. The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
% \end{figure}

% \newpage
